### 2017-02-27 2017mergingmerges/section/05-Evaluation.tex
Put pipelined combinators graph first
S5.2 leave talk of result size for now

In practice, the likelihood of fusion suceeding depends on the particular dataflow network being used, as well as the form of the processes in that network. For fusion of pipelines of standard combinators such as @map@, @fold@, @filter@, @scan@ and so on, fusion always succeeds. The process implementations of each of these combinators only pull one element at a time from their source streams, before pushing the result to the output stream, so there is no possiblity of deadlock. Deadlock can only happen when multiple streams fan-in to a process with multiple inputs, such as with @merge@. When the dataflow network has a single output stream then we use the method of starting from the process closes to the output stream, walking to towards the input streams, and fusing in successive processes as they occur. This allows the interleaving of the intermediate fused process to be dominated by the consumers, rather than producers, as consumers are more likely to have multiple input channels where need to be synchronized. In the worst case the fallback approach is to try all possible orderings of processes to fuse, assuming the client programmer is willing to wait for the search to complete.

@@

\TODO{Discuss the performance of the system. Why this is a good idea. Give graphs that show the number of states is sensible when more processes are fused in.}

@@

\begin{figure}

% I thought these two looked nicer separately, but let me know if you want them in a single graph
\begin{minipage}[t]{0.4\textwidth}
\begin{tikzpicture}
\begin{axis}[
	ylabel=Process states,
	xlabel=Split combinators,
  ymin=0, ymax=100,
	enlargelimits=0.01,
	ybar interval=0.7,
  width=7.5cm, height=4.5cm,
	legend pos=north west,
]
\addplot coordinates {(1,9) (2,27) (3,40) (4,53) (5,66) (6,79) (7,92)
  % Last bar doesn't show for some reason, so need to add a dummy value for the next one
    (8,1) };

@@

\legend{With merge, Without merge};

\end{axis}
\end{tikzpicture}
\end{minipage}
\begin{minipage}[t]{0.10\textwidth}
\quad
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{tikzpicture}
\begin{axis}[
% Hide the label on the second graph
%	ylabel=Process states,
	xlabel=Pipelined combinators,
  ymin=0, ymax=100,
	enlargelimits=0.01,
	ybar interval=0.7,
  width=7.5cm, height=4.5cm,
	legend pos=north west,
]
\addplot coordinates {(1,9) (2,17) (3,27) (4,37) (5,47) (6,57) (7,67)   (8,1) };

\legend{With merge, Without merge};

\end{axis}
\end{tikzpicture}
\end{minipage}

@@

\begin{figure}

>>>>

In practice, the likelihood of fusion suceeding depends on the particular dataflow network being used, as well as the form of the processes in that network. For fusion of pipelines of standard combinators such as @map@, @fold@, @filter@, @scan@ and so on, fusion always succeeds. The process implementations of each of these combinators only pull one element at a time from their source streams, before pushing the result to the output stream, so there is no possiblity of deadlock. Deadlock can only happen when multiple streams fan-in to a process with multiple inputs, such as with @merge@. When the dataflow network has a single output stream then we use the method of starting from the process closes to the output stream, walking to towards the input streams, and fusing in successive processes as they occur. This allows the interleaving of the intermediate fused process to be dominated by the consumers, rather than producers, as consumers are more likely to have multiple input channels which need to be synchronized. In the worst case the fallback approach is to try all possible orderings of processes to fuse, assuming the client programmer is willing to wait for the search to complete.

@@

As with fusion transformation system, we must be careful that the size of the result code does not become too large when more and more processes are fused together. The left of Fig.\ref{fig:bench:outputsize} shows the maximum number of output states in the result when a particular number of processes are fused together in a pipelined-manner. To produce this graph we programmatically generated dataflow networks for \emph{all possible} pipelined combinations of the @map@, @filter@, @scan@, @group@ and @merge@ combinators, and tried all possible fusion orders consiting of adjacent pairs of processes. The @merge@ combinator itself has two inputs, so only works at the very start of the pipeline --- we present result for pipelines with and without a @merge@ at the start. The right of Fig.\ref{fig:bench:outputsize} shows the number of states in the result when the various combinations of combinators are fused in parallel, for example, we might have a @map@ and a @filter@ processing the same input stream. In both cases the number of states in the result process grows linearly with the number of processes. In all combinations, with up to 7 processes there are less than 100 states in the result process.

The size of the result process is roughly what one would get when inlining the definitions of each of the original source processes. This is common with other systems based on inlining and/or template meta-programming, and is not prohibitive.

On the other hand, Fig.\ref{fig:bench:exponential} shows the results for a pathological case where the size of the output program is exponential in the number if input processes. The source dataflow networks consists of N merge processes, N+1 input streams, and a single output stream. The output of each merge process is the input of the next, forming a chain of merges. In source notation the network for N = 3 is @sOut = merge sIn1 (merge sIn2 (merge sIn3 sIn4))@.

When fusing two processes the fusion algorithm essentially compares every state in the first process with every state in the second, computing a cross product. During the fusion transform, as states in the result process are generated they are added to a finite map --- the @instrs@ field of the process definition. The use of the finite map ensures that identical states are always combined, but genuinly different states always make it into the result.

In the worst case fusion of two processes produces O($n*m$) different states, where $n$ and $m$ are the number of states in each. If we assume the two processes have about the same number of states then this is O($n^2$). Fusing the next process into this result yields O($n^3$), so overall the worst case number of states in the result will be O($n^k$), where $k$ is the number of processes fused.

In the particular case of @merge@, the implementation has two occurrences of the @push@ instruction. During fusion, the states for the consuming process are inlined at each occurrence of @push@. These states are legitimately different because at each occurence of @push@ the input channels of the merge process are in different channel states, and these channel states are included in the overall process state.

@@

% -----------------------------------------------------------------------------
\begin{figure}
\begin{minipage}[t]{0.10\textwidth}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{tikzpicture}
\begin{axis}[
% Hide the label on the second graph
	ylabel=Process states,
	xlabel=Pipelined combinators,
  ymin=0, ymax=100,
	enlargelimits=0.01,
	ybar interval=0.7,
  width=7.5cm, height=4.5cm,
	legend pos=north west,
]
\addplot coordinates {(1,9) (2,17) (3,27) (4,37) (5,47) (6,57) (7,67)   (8,1) };
\addplot coordinates {(1,4) (2,9) (3,14) (4,19) (5,24) (6,29) (7,34)    (8,1) };
\legend{With merge, Without merge};
\end{axis}
\end{tikzpicture}
\end{minipage}
\quad
\quad
\quad
\begin{minipage}[t]{0.4\textwidth}
\begin{tikzpicture}
\begin{axis}[
%       ylabel=Process states,
        xlabel=Parallel combinators,
  ymin=0, ymax=100,
        enlargelimits=0.01,
        ybar interval=0.7,
  width=7.5cm, height=4.5cm,
        legend pos=north west,
]
\addplot coordinates {(1,9) (2,27) (3,40) (4,53) (5,66) (6,79) (7,92)
  % Last bar doesn't show for some reason, so need to add a dummy value for the next one
    (8,1) };

@@

\legend{With merge, Without merge};
\end{axis}
\end{tikzpicture}
\end{minipage}

@@

% -----------------------------------------------------------------------------
\begin{figure}
